---
title: Assessing the association cell-specific DNAm and atopic dermatitis 
author: Thomas Battram
output:
  bookdown::html_document2:
    toc: true
    theme: united
params:
  man: ""
  qq: ""
  summ: ""
  workflow: ""
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}

## pkgs
suppressWarnings(suppressPackageStartupMessages({
library(bookdown) # extra markdown options
library(knitr) # ...
library(kableExtra) # extra nice tables
library(tidyverse) # tidy code and data
library(ewaff) # For easy EWAS plots
library(cowplot) # plotting figures on a grid
library(RColorBrewer) # for plotting colours
library(usefunc) # own package of useful functions
}))

knitr::opts_chunk$set(echo = FALSE, 
           			  warning = FALSE, 
            		  message = TRUE, 
            		  cache = FALSE, 
            		  dpi = 300 
            		  )

## params tester
# params <- list(man = "results/celldmc-tca-manhattans.png", qq = "results/celldmc-tca-qqs.png", summ = "results/summary-of-results.RData", workflow = "~/projects/ad-cse/ewas/report/cell-spec-workflow.drawio.pdf")

# load summ data
summ <- new_load(params$summ)

```

## Introduction

There is little evidence for a strong association between DNAm measured at birth and childhood AD. There have been other papers suggesting there are associations when measuring DNAm at later time points. Previous studies have used bulk tissue, but changes in DNAm may have cell-specific consequences. The most recent AD GWAS found enrichments for hits in DNAse hypersensitive regions amongst B cells, T cells and NK cells, suggesting cell-specific regulation in blood contributes to disease development. Recently, methods have been developed aiming to estimate associations between cell-specific DNAm and traits using DNAm measured in bulk tissue and a reference. We used the methods, CellDMC, TCA, and omicWAS to estimate cell-specific associations between AD and DNAm. Workflow is shown in __Figure \@ref(fig:workflow)__

(ref:workflow-caption) Workflow. Discovery: P = 1.42x10-8 (1x10-7 / 7), Replication: P = 0.05 / N hits AND effect estimate in same direction.

```{r workflow, results = "asis", fig.cap = '(ref:workflow-caption)'}
include_graphics(params$workflow)
```

## Methods

We used ARIES 15up samples and defined teenage AD using parental questionnaire data, where the parent was asked whether their child had eczema within the last 12 months. This question was asked at 3 timepoints: 16y 6m, 13y 10m, 13y 1m. If "Yes" was answered to any of these questions they were deemed a case, regardless of whether they had been to a doctor. This gave us `r summ$samplesizes$N_case` cases and `r summ$samplesizes$N_controls` controls after QC.

DNAm data was taken from the 15up timepoint and measured using the EPIC array. `r summ$n_cpgs` CpG sites remained after QC.

### Models

1. DNAm ~ AD + sex + age + surrogate variables 
2. DNAm ~ AD + sex + surrogate variables + cell counts

Other variables needed? - Genetic principal components weren't used because lots of individuals would be removed as they don't have genotype data

Note: Model 2 is only for the conventional EWAS

## Results

```{r summ-data-res}
n_hits <- length(unique(summ$initial_hits$CpG))
n_cells <- length(unique(summ$initial_hits[["Cell type"]]))

omicwas_p_thresh <- 0.05 / n_hits
omic_rep_cpgs <- summ$initial_hits %>%
  dplyr::filter(Method == "omicWAS") %>%
  dplyr::filter(P < omicwas_p_thresh) %>%
  pull("CpG")

## ewaff hits
cc_hits <- summ$ewaff_hits %>%
  dplyr::filter(Model != "unadjusted") %>%
  dplyr::filter(P < 1e-7) %>%
  pull(CpG)

nocc_hits <- summ$ewaff_hits %>%
  dplyr::filter(Model == "unadjusted") %>%
  dplyr::filter(P < 1e-7) %>%
  pull(CpG)  
```

There were `r n_hits` sites within `r n_cells` cells that were identified by CellDMC and TCA. Of these, `r length(omic_rep_cpgs)` replicated when using omicWAS __Table \@ref(tab:initial-hits)__. In the conventional EWAS, `r length(cc_hits)` CpG sites associated with AD when adjusting for cell counts and `r length(nocc_hits)` associated when not adjusting for cell counts.

```{r cs-qqs, results = "asis", fig.cap = "CellDMC and TCA QQ plots"}
include_graphics(params$qq)
```

```{r cs-mans, results = "asis", fig.cap = "CellDMC and TCA manhattan plots"}
include_graphics(params$man)
```

```{r initial-hits}
kbl(summ$initial_hits, caption = "Association between AD and cell-specific DNAm") %>%
    kable_styling()
```
